\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{caption}
\usepackage{subcaption}
\doublespacing
\title{Distributional Impacts of Electricity Rate Design Reform }
\author{
Madalsa Singh\thanks{Corresponding author. Email: [email]} \\
\textit{[Your Institution]} \\
\and
Patricia Hidalgo-Gonzalez \\
\textit{[Institution]} \\
\and
Rajni Bansal \\
\textit{[Institution]}\\
\and
Ranjit Deshmukh \\
\textit{[Institution]}
}
\date{\today}
\begin{document}
\maketitle
\begin{abstract}
This paper examines the distributional impacts of electricity rate design reforms in California's investor-owned utilities. We analyze how shifts from volumetric to fixed-charge pricing structures, combined with policy interventions such as wildfire cost removal and return-on-equity reductions, affect residential customers across income levels, consumption patterns, and climate zones. Using 4,898 synthetic building profiles representative of San Diego Gas \& Electric's (SDGE) 1.3 million residential customers, we design revenue-neutral rate structures and calculate household-level bill impacts under 40 policy scenarios. We extend the analysis to account for distributed energy resource adoption---rooftop solar, battery storage, electric vehicles, and heat pumps---by assigning technologies to buildings using survey-derived adoption probabilities and computing bills under California's Net Billing Tariff (NEM~3.0). Our findings reveal systematic distributional effects: fixed charges redistribute costs from high-consumption to low-consumption households, with impacts varying significantly by income level and geographic location. Policy interventions to remove wildfire costs and reduce utility returns on equity reduce bills across all customer groups but provide larger percentage savings to lower-income households. The interaction between rate design and technology adoption is substantial: fixed charges reduce the bill savings from solar self-consumption, while time-of-use rate spreads influence the value of battery arbitrage.

\textbf{Keywords:} Electricity pricing, rate design, fixed charges, distributional equity, California, time-of-use rates, net billing, distributed energy resources
\end{abstract}

\section{Introduction}
California's electricity sector faces mounting financial pressures from wildfire mitigation costs, grid modernization investments, and the transition to renewable energy. The most recent cost increases have translated into residential electricity rates in the three investor-owned utilities to be among the highest in the continental United States, with San Diego Gas \& Electric (SDGE) customers facing average rates exceeding \$0.50 per kilowatt-hour (kWh) \citep{[citation needed]}. This high-rate environment has intensified policy debates over electricity rate design, particularly regarding the balance between volumetric charges (based on consumption) and fixed monthly charges.

In 2024, the California Public Utilities Commission (CPUC) authorized investor-owned utilities to implement income-graduated fixed charges (IGFCs) as part of broader rate reform efforts under Rulemaking R.22-07-005, pursuant to Assembly Bill 205 (2022). The CPUC's Decision D.24-05-028 adopted a three-tier structure: \$24.15/month for general residential customers, \$12.08/month for FERA-eligible customers, and \$6/month for CARE-eligible customers, with corresponding reductions in volumetric rates of 5--7 cents/kWh \citep{CPUC_D2405028}. SDGE's implementation (effective December 2025) sets fixed charges at \$6.11/month for CARE and \$24.60/month for non-CARE customers.

Proponents argue that fixed charges improve utility cost recovery, reduce rate volatility, and better align charges with the cost structure of distribution systems. The IOUs initially proposed substantially higher fixed charges---\$53 (PG\&E), \$74 (SDGE), and \$49 (SCE) per month---contending that volumetric-only rates overcharge high-consumption customers for infrastructure that benefits all grid-connected households \citep{JointIOUs_R2207005}. Critics contend that fixed charges reduce incentives for energy conservation, disproportionately burden low-consumption households, and undermine distributed energy resource adoption. The Greenlining Institute characterized the adopted structure as ``not an income-graduated rate structure [but] a flat rate charge that poses the same regressivity red flags that it was supposed to address'' \citep{Greenlining2024}. TURN and NRDC's joint testimony proposed a more steeply graduated alternative---approximately \$5/month for CARE customers, \$42 for middle-income households, and \$62 for high-income households---arguing that a three-tier system with modest graduation would leave low-usage lower-middle-class customers worse off while reducing bills for higher-consumption wealthy households \citep{NRDC_TURN_01}. The California Environmental Justice Alliance (CEJA) similarly criticized the decision for ``lumping households that make \$65,000 into the highest income tier along with millionaires'' \citep{CEJA2024}.

This study contributes to the electricity rate design literature by providing granular, building-level analysis of how different rate structures affect customer bills across the distribution of income, consumption, and geography. While previous studies have examined rate design impacts using aggregate or representative customer archetypes \citep{[citations]}, our approach uses detailed synthetic building profiles to capture heterogeneity within customer populations. We analyze not only the current fixed charge policy but also alternative rate designs incorporating wildfire cost removal and utility return-on-equity reductions---policy interventions under active consideration in California.

A critical gap in existing rate design analyses is the omission of distributed energy resources (DERs). Approximately 15--20\% of SDGE residential customers have adopted rooftop solar PV, and battery storage, electric vehicles, and heat pumps are increasingly prevalent. These technologies fundamentally alter households' net consumption profiles and, consequently, how rate design changes affect their bills. We address this by supplementing the baseline building stock with empirically-grounded technology adoption assignments derived from a survey of California residential customers, and by modeling bill impacts under California's Net Billing Tariff (NEM~3.0), which compensates solar exports at hourly avoided-cost rates rather than retail rates.

Our analysis focuses on SDGE's service territory as a case study, with plans to extend the methodology to all California investor-owned utilities (Pacific Gas \& Electric and Southern California Edison). SDGE serves approximately 1.3 million residential customers across three distinct climate zones (California Energy Commission zones 7, 10, and 14), providing meaningful variation in consumption patterns and building characteristics.

\section{Methods}

\subsection{Data Sources}

\subsubsection{Building Profiles}
We obtained residential building energy consumption profiles from the National Renewable Energy Laboratory (NREL) ResStock dataset \citep{[ResStock citation]}, which provides synthetic hourly electricity consumption for representative building stocks across the United States. We extracted 4,898 building profiles located in San Diego County, identified via census Public Use Microdata Area (PUMA) codes corresponding to SDGE's service territory, from the California Baseline dataset. Each building profile includes:
\begin{itemize}
    \item Hourly electricity consumption at 15-minute resolution for a typical meteorological year (8,760 hours)
    \item Building characteristics (square footage, vintage, construction type, HVAC systems)
    \item Income classification based on area median income (AMI) thresholds
    \item California Energy Commission (CEC) climate zone designation
    \item A uniform statistical weighting factor of 252.3, representing the number of actual households each sample building represents
\end{itemize}

The 4,898 sample buildings represent approximately 1.24 million weighted households---93.4\% of SDGE's 1.32 million residential customers. The sample spans 23 PUMAs across three primary CEC climate zones: CZ~7 (coastal San Diego, 71.2\% of sample), CZ~10 (inland valleys, 27.7\%), and CZ~14 (mountain/desert, 0.8\%), with a small number of buildings in adjacent zones CZ~6 and CZ~15.

\subsubsection{Sample Representativeness}
\label{sec:representativeness}

Table~\ref{tab:sample_comparison} compares key characteristics of our simulated building sample against actual SDGE service territory data and San Diego County census statistics. The sample closely reproduces the observed housing stock composition: single-family detached homes constitute 50.6\% of the sample versus approximately 51\% in ACS data, multi-family units represent 35.6\% versus approximately 35\%, and the homeownership rate of 53.0\% aligns with the county-wide estimate of 54\%.

The primary discrepancy between the simulated sample and actual SDGE data is in average electricity consumption. Before RASS calibration, the raw ResStock sample mean for SDGE territory is 7,021~kWh/year, substantially exceeding observed consumption. After applying RASS calibration (described in Section~\ref{sec:rass}), the sample mean is reduced to 5,309~kWh per customer per year---a meaningful improvement, but still exceeding SDGE's reported average of 3,634~kWh by approximately 46\%. This remaining discrepancy arises primarily because SDGE's average includes very low-consumption accounts (vacant units, partial-year occupancy, common-area meters) that are not represented in ResStock's occupied-building sample. The RASS calibration itself is effective at correcting the building physics bias: scaling factors below 1.0 for CZ~7 (0.713) and CZ~10 (0.841) reflect ResStock's tendency to overestimate cooling loads in mild coastal climates, while factors above 1.0 for CZ~15 (1.417) correct underestimates in desert zones. Despite the remaining level difference after RASS scaling, the \textit{relative} consumption patterns across climate zones, housing types, and income groups are well-preserved, which is the relevant feature for our distributional analysis of rate design impacts.

The CARE eligibility rate in our sample (25.9\% based on income classification) is modestly below SDGE's reported 28.1\% participation rate. This difference is within the range expected given that CARE eligibility depends on household size and income thresholds that are approximated by ResStock's AMI-based classifications.

\input{table_sample_comparison}

\subsubsection{RASS Calibration}
\label{sec:rass}

We calibrated the ResStock consumption profiles against the 2019 Residential Appliance Saturation Study (RASS) \citep{[RASS citation]}, which provides empirical estimates of residential electricity consumption by climate zone for California households.

\textbf{Climate Zone Scaling Factors.} For each CEC climate zone, we computed the ratio of RASS survey mean consumption to ResStock simulated mean consumption. The resulting scaling factors were applied multiplicatively to each building's hourly load profile \textit{before} any bill calculations, ensuring that fixed charges (which are per-household, not per-kWh) are unaffected by the consumption adjustment:
\begin{equation}
Q_{i,h}^{\text{scaled}} = Q_{i,h}^{\text{raw}} \times s_{z(i)}
\end{equation}
where $Q_{i,h}$ is building $i$'s consumption in hour $h$ and $s_{z(i)}$ is the scaling factor for building $i$'s climate zone $z$. The scaling factors range from 0.713 (CZ~7, coastal) to 1.417 (CZ~15, desert), reflecting ResStock's tendency to overestimate cooling loads in mild coastal climates and underestimate them in hot desert zones. To mitigate the influence of extreme outliers, we winsorized consumption values at the 1st and 99th percentiles within each climate zone before computing scaling factors.

\textbf{Income-Energy Relationship.} We estimated log-log regressions for each SDGE climate zone to quantify the elasticity of electricity consumption with respect to household income:
\begin{equation}
\log(E_{i}) = \beta_0 + \beta_1 \log(I_{i}) + \epsilon_i
\end{equation}
where $E_i$ is annual electricity consumption (kWh) and $I_i$ is household income. The estimated elasticities were:
\begin{itemize}
    \item Climate Zone 7 (Coastal): $\beta_1 = 0.277$ (mean consumption: 4,824 kWh/year)
    \item Climate Zone 10 (Inland): $\beta_1 = 0.270$ (mean consumption: 6,456 kWh/year)
    \item Climate Zone 14 (Mountain): $\beta_1 = 0.154$ (mean consumption: 7,500 kWh/year)
\end{itemize}
These elasticity estimates indicate that a 1\% increase in income is associated with approximately 0.15--0.28\% increase in electricity consumption, with coastal areas (CZ~7) showing higher income sensitivity than inland regions (CZ~14), where climate-driven cooling loads dominate consumption patterns regardless of income.

\subsubsection{Rate Structures}
Current SDGE residential rate structures were obtained from the utility's published tariff schedules (October 2025). We analyzed four actual rate schedules:
\begin{enumerate}
    \item \textbf{DR}: Tiered rate (pre-fixed charge)
    \item \textbf{DR-F}: Tiered rate with income-graduated fixed charges
    \item \textbf{TOU-DR}: Time-of-use rate (pre-fixed charge)---serves 68\% of customers
    \item \textbf{TOU-DR-F}: Time-of-use rate with fixed charges (\$6.11/month CARE, \$24.60/month non-CARE)
\end{enumerate}
We use TOU-DR as the primary benchmark, as it represents the majority of SDGE residential customers. The TOU-DR tariff has summer peak rates of \$0.60/kWh, summer off-peak of \$0.45/kWh, and a baseline credit of \$0.11/kWh applied to consumption within the baseline allowance. CARE customers receive a 37\% discount on volumetric rates. These actual tariff rates serve as the starting point for the designed rate scenarios described in Section~\ref{sec:tariff_validation}.

\subsubsection{Utility Financial Data}
Revenue requirements, rate base, and cost components were obtained from SDGE's General Rate Case filings and CPUC decisions \citep{[GRC citation]}. Key parameters include:
\begin{itemize}
    \item Total residential revenue requirement: \$1.562 billion (from SDGE's 2024 General Rate Case filing, representing the authorized revenue the utility is permitted to collect from residential customers---not a modeled quantity)
    \item Total residential sales: 4,810 GWh
    \item CARE customers: 372,135 (28.1\% of total)
    \item Non-CARE customers: 951,477 (71.9\% of total)
    \item Authorized return on equity: 10.22\%
    \item Rate base: \$13.59 billion (utility-wide)
    \item Capital structure: 52\% equity / 48\% debt
    \item Wildfire fund recovery: \$414 million (utility-wide)
    \item Transmission costs: \$685 million (utility-wide)
    \item Distribution costs: \$1.722 billion (utility-wide)
\end{itemize}
Cost allocation to the residential class was calculated proportionally to residential revenue share of total utility revenue (36.9\%), yielding residential wildfire costs of \$153M, residential transmission costs of \$253M, and residential distribution costs of \$636M.

\subsubsection{Energy Export Credit (EEC) Data}
\label{sec:eec_data}

For net billing calculations, we constructed an 8,760-hour table of Energy Export Credit rates for all three California IOUs from utility-published pricing files for the 2025 interconnection vintage:
\begin{itemize}
    \item \textbf{PG\&E}: Extracted from the ``Solar Billing Plan EEC Values Price Sheet 2025'' PDF, which provides separate ``Produced'' and ``Delivered'' components by hour-of-day, month, and day type (weekday/weekend).
    \item \textbf{SCE}: Parsed from the ``NBT25 MIDAS File'' Excel workbook, with Produced (code USCA-XXSC) and Delivered (code USCA-SCXX) components separated.
    \item \textbf{SDGE}: Parsed from the ``LY2025 NBT Pricing Upload MIDAS'' CSV, which provides a single combined rate.
\end{itemize}
MIDAS files use an 8am-to-8am day cycle, which we remapped to calendar hours. Holidays follow the NERC schedule plus California-observed holidays. The resulting table provides hourly EEC rates that vary substantially across hours and seasons, with near-zero midday values reflecting high solar penetration and higher evening values reflecting scarcity pricing.

\subsection{Rate Design Methodology}

We developed a revenue-neutral rate design framework that maintains total utility revenue while varying the allocation between fixed and volumetric charges. The methodology ensures that scenarios with different fixed charge levels but the same policy parameters (wildfire treatment, ROE adjustment) collect identical total revenue---a critical constraint for valid distributional comparisons.

\subsubsection{Time-of-Use Consumption Weights}

A key element of our rate design approach is the calculation of consumption-weighted time-of-use (TOU) period weights directly from the building sample's hourly load profiles. We classified each hour of consumption into six TOU periods based on SDGE's rate schedule:
\begin{itemize}
    \item \textbf{Summer} (June--October): Peak (4--9pm, 14.4\%), Off-peak (10pm--6am, 12.9\%), Super off-peak (6am--4pm \& 9--10pm, 20.6\%)
    \item \textbf{Winter} (November--May): Peak (4--9pm, 14.2\%), Off-peak (10pm--6am, 15.1\%), Super off-peak (6am--4pm \& 9--10pm, 22.8\%)
\end{itemize}
These empirical weights, computed from the 4,898-building sample, ensure that designed rates yield the target revenue when applied to the sample's actual consumption patterns. The weights were saved and reused consistently across all rate scenarios.

\subsubsection{Revenue-Neutral Rate Calculation}

For each scenario, we first determined the target revenue requirement:
\begin{equation}
R_{\text{target}} = R_{\text{base}} - \mathbb{1}_{\text{WF}} \cdot C_{\text{wildfire}} - \Delta_{\text{ROE}} \cdot B_{\text{rate}} \cdot \phi_{\text{eq}} \cdot s_{\text{res}}
\end{equation}
where $R_{\text{base}} = \$1.562$B is the baseline residential revenue, $C_{\text{wildfire}}$ is the residential share of wildfire costs (\$153M), $\Delta_{\text{ROE}}$ is the percentage-point ROE reduction, $B_{\text{rate}}$ is the utility rate base (\$13.59B), $\phi_{\text{eq}} = 0.52$ is the equity share of the utility's capital structure (since ROE applies only to the equity portion, not debt), and $s_{\text{res}} = 0.369$ is the residential revenue share. SDGE's authorized ROE of 10.22\% and its 52\%/48\% equity-debt capital structure imply that a 1 percentage-point ROE reduction lowers the residential revenue requirement by approximately \$26M.

Fixed charge revenue was computed based on transmission and distribution (T\&D) cost recovery:
\begin{equation}
R_{\text{fixed}} = (C_{\text{T}} + C_{\text{D}}) \times \frac{\alpha}{100}
\end{equation}
where $C_{\text{T}}$ and $C_{\text{D}}$ are residential transmission and distribution costs, and $\alpha \in \{0, 25, 50, 75\}$ is the percentage of T\&D costs recovered through fixed charges.

Income-graduated monthly fixed charges were set as:
\begin{equation}
F_{\text{non-CARE}} = \frac{R_{\text{fixed}}}{12 \times (N_{\text{non-CARE}} + \beta \cdot N_{\text{CARE}})}, \qquad F_{\text{CARE}} = \beta \cdot F_{\text{non-CARE}}
\end{equation}
where $\beta = 0.4$ is the CARE discount factor, reflecting the policy that CARE customers pay 40\% of the standard fixed charge.

The target average volumetric rate was then:
\begin{equation}
\bar{r}_{\text{target}} = \frac{R_{\text{target}} - R_{\text{fixed}}}{Q_{\text{total}}}
\end{equation}
where $Q_{\text{total}} = 4{,}810$ GWh is total residential sales. Finally, all six TOU rates were scaled proportionally from the baseline rate structure:
\begin{equation}
r_{p,s}^{\text{new}} = r_{p,s}^{\text{baseline}} \times \frac{\bar{r}_{\text{target}}}{\sum_{p,s} r_{p,s}^{\text{baseline}} \cdot w_{p,s}}
\end{equation}
where $r_{p,s}^{\text{baseline}}$ is the current rate for period $p$ in season $s$, and $w_{p,s}$ is the empirical consumption weight.

This proportional scaling preserves the relative price signals across TOU periods (i.e., the peak-to-off-peak ratio remains constant) while adjusting the overall rate level to achieve revenue neutrality. We verified that for all scenarios with the same policy parameters, the coefficient of variation of total revenue across fixed charge levels was exactly 0\%.

\subsection{Scenario Design}

We analyzed a factorial design combining three policy dimensions:
\begin{enumerate}
    \item \textbf{Fixed charge levels}: 0\%, 25\%, 50\%, 75\%, 100\% of T\&D costs recovered through fixed charges (5 levels)
    \item \textbf{Wildfire cost treatment}: Include or remove \$153 million in residential wildfire mitigation costs (2 levels)
    \item \textbf{Return on equity adjustment}: Baseline (10.22\%), or reduce by 0.5, 1.0, or 1.5 percentage points (4 levels)
\end{enumerate}
The full factorial produces $5 \times 2 \times 4 = 40$ rate scenarios. For the primary analysis, we focus on a subset of eight scenarios that isolate each policy lever and their key combinations, plus the two actual SDGE tariff rates for benchmarking:
\begin{enumerate}
    \item \textbf{TOU-DR}: Actual SDGE time-of-use rate (no fixed charges)---the benchmark
    \item \textbf{TOU-DR-F}: Actual SDGE rate with income-graduated fixed charges (\$6.11/\$24.60 per month)
    \item \textbf{F0\_WF0\_ROE0}: Designed status quo (all volumetric, revenue-neutral baseline)
    \item \textbf{F50\_WF0\_ROE0}: 50\% of T\&D costs to fixed charges
    \item \textbf{F100\_WF0\_ROE0}: 100\% of T\&D costs to fixed charges
    \item \textbf{F0\_WF1\_ROE0}: Remove wildfire costs (all volumetric)
    \item \textbf{F0\_WF0\_ROE1.0}: 1 percentage-point ROE reduction
    \item \textbf{F50\_WF1\_ROE1.0}: Combined scenario (50\% fixed + wildfire removal + 1pp ROE cut)
\end{enumerate}
This focused set allows us to isolate the distributional effect of each policy lever individually and in combination, while the full 40-scenario grid supports sensitivity analysis.

\subsection{Bill Calculation}

Total annual bills were calculated as the sum of fixed and volumetric components, accounting for the temporal structure of electricity consumption and rate schedules.

\subsubsection{Fixed Charge Component}
Monthly fixed charges were applied based on CARE eligibility:
\begin{equation}
C_{\text{fixed}} = 12 \times F_{j(i)}
\end{equation}
where $F_{j(i)} \in \{F_{\text{CARE}}, F_{\text{non-CARE}}\}$ depends on customer $i$'s income classification.

\subsubsection{Volumetric Charge Component}
For time-of-use rates, volumetric charges were computed by matching each hour's (RASS-scaled) consumption to the corresponding rate:
\begin{equation}
C_{\text{vol}} = \sum_{h=1}^{8760} Q_{i,h}^{\text{scaled}} \times r_{p(h),s(h)}
\end{equation}
where $p(h)$ and $s(h)$ map hour $h$ to its TOU period and season.

\subsubsection{Total Bill and Revenue Neutrality}
The complete annual bill is:
\begin{equation}
B_{i} = C_{\text{fixed},i} + C_{\text{vol},i}
\end{equation}

Revenue neutrality was verified by computing weighted total revenue across the sample:
\begin{equation}
R_{\text{total}} = \sum_{i=1}^{N} B_{i} \times w_i
\end{equation}
and confirming that scenarios with identical policy parameters yield identical total revenue regardless of fixed charge level. Revenue deviations were required to be within $\pm$1\% of the target for all scenarios.

\subsubsection{Reconciling Sample and Utility Revenue}
\label{sec:revenue_reconciliation}

Because the simulated sample's average consumption (5,309~kWh) exceeds SDGE's reported average (3,634~kWh) by approximately 46\%, applying SDGE's published tariff rates to our sample generates total revenue that exceeds the utility's filed residential revenue requirement of \$1.562B. Importantly, the \$1.562B figure is not a modeled quantity---it is SDGE's authorized revenue from its General Rate Case filing, representing the actual amount the utility is permitted to collect from residential customers. The 46\% consumption discrepancy therefore does \textit{not} inflate our rate designs or revenue targets by 1.46$\times$; rather, it means that if one were to naively extrapolate sample-average bills to the full 1.32~million customer population, the implied aggregate revenue would exceed the filed amount.

Our rate design avoids this problem by using SDGE's actual total residential sales (4,810~GWh) and customer counts in the revenue-neutrality equations, ensuring that the \textit{designed rates} (the \$/kWh values) are calibrated to SDGE's real economics. The rate \textit{structure}---the per-kWh charges that determine how bills vary across customers---comes directly from the actual SDGE tariff, not from dividing filed revenue by simulated sales. When these rates are applied to the simulated buildings for bill calculation, the resulting per-building bills reflect what households with those consumption profiles would actually pay under each rate scenario. The distributional analysis---comparing bill changes across income groups, consumption levels, and climate zones---is valid because it depends on the \textit{relative} structure of rates and consumption, not on aggregate revenue matching. We discuss this limitation further in Section~\ref{sec:limitations}.

\subsection{Distributional Analysis}

We analyzed bill impacts across three dimensions:
\begin{enumerate}
    \item \textbf{Income level}: Low (CARE-eligible), Medium, High based on AMI thresholds
    \item \textbf{Annual consumption}: Deciles from lowest to highest consumers
    \item \textbf{Climate zone}: CZ~7 (coastal San Diego), CZ~10 (inland valleys), CZ~14 (mountain/desert)
\end{enumerate}

For each scenario relative to the TOU-DR baseline (no fixed charge, no policy intervention), we calculated percentage and absolute bill changes:
\begin{equation}
\Delta B_i^{\%} = \frac{B_i^{\text{scenario}} - B_i^{\text{baseline}}}{B_i^{\text{baseline}}} \times 100\%, \qquad
\Delta B_i^{\$} = B_i^{\text{scenario}} - B_i^{\text{baseline}}
\end{equation}

Results are presented as distributions (median, interquartile range) within each customer segment to capture within-group heterogeneity.

\subsection{Tariff-Based Rate Validation}
\label{sec:tariff_validation}

To benchmark our designed rate scenarios against actual customer experience, we implemented a parallel bill calculation using SDGE's published tariff rates. The actual TOU-DR tariff has volumetric rates substantially higher than the revenue-divided average (e.g., summer peak of \$0.60/kWh vs.\ the revenue-average of \$0.3247/kWh), because the tariff rates incorporate baseline credits ($\sim$\$0.11/kWh applied to within-baseline consumption) and CARE discounts (37\% on volumetric rates) that reduce effective rates below the posted tariff.

We validated the designed rate scenarios by confirming that the actual TOU-DR-F tariff---which represents the CPUC's fixed charge ruling at approximately 35\% of T\&D costs in fixed charges---is reproduced by our rate design framework to within $<$1\% error on all six TOU period rates and the income-graduated fixed charges (\$6.11 CARE, \$24.60 non-CARE). This cross-validation ensures that the designed scenarios (which vary fixed charge levels parametrically) are anchored to the real tariff structure.

For baseline bill comparisons, we compute bills under both the actual SDGE tariff (using the full bill calculator with baseline credits, tier structures, and CARE discounts) and the designed revenue-neutral rates. We report bills under the actual TOU-DR tariff as the primary benchmark, with designed scenarios providing the counterfactual rate structures for policy analysis.

\subsection{Net Billing (NEM 3.0)}
\label{sec:net_billing}

For buildings with rooftop solar PV, we calculate bills under California's Net Billing Tariff (also known as the Solar Billing Plan or NEM~3.0), which replaced net energy metering in April 2023. Under net billing, import and export are settled asymmetrically within each hour:
\begin{itemize}
    \item \textbf{Import hours} ($Q_{i,h}^{\text{net}} > 0$): The customer pays the full retail TOU rate on net consumption.
    \item \textbf{Export hours} ($Q_{i,h}^{\text{net}} < 0$): The customer earns an Energy Export Credit (EEC) on net exports, valued at the hourly avoided cost rate.
\end{itemize}

The EEC rates are published by each utility and vary by hour, reflecting the time-varying avoided cost of generation, capacity, transmission, and distribution. For SDGE, we constructed an 8,760-hour EEC table from the utility's MIDAS pricing files for the 2025 interconnection vintage. Mean annual EEC rates are approximately \$0.06/kWh for SDGE, \$0.09/kWh for PG\&E, and \$0.08/kWh for SCE---substantially below retail rates of \$0.45--0.60/kWh, creating a strong incentive for self-consumption over export.

Monthly export credits offset import charges, with excess credits rolling forward. At annual true-up, remaining credits are forfeited (valued at \$0). The annual net billing calculation is:
\begin{equation}
B_i^{\text{NB}} = \sum_{m=1}^{12} \max\left(0,\; C_{\text{import},m} - C_{\text{export},m} + F_m - \text{Credits}_{m-1}\right)
\end{equation}
where $C_{\text{import},m}$ and $C_{\text{export},m}$ are the monthly import charges and export credits, $F_m$ is the monthly fixed charge, and $\text{Credits}_{m-1}$ is the carried-forward credit from the previous month.

\subsection{Technology Adoption and Load Modification}
\label{sec:tech_adoption_main}

The ResStock baseline profiles represent static load profiles without distributed energy resources increasingly prevalent in SDGE's territory. To capture the interaction between rate design and technology adoption, we assign four technologies---solar PV, battery energy storage, electric vehicles, and heat pumps---to buildings in the sample using survey-derived adoption probabilities calibrated to current administrative penetration rates.

\subsubsection{Adoption Assignment}

We use a supplemental adoption approach that preserves ResStock's existing assignments (6.2\% solar PV, 8.3\% heat pumps) and adds additional adopters to reach current SDGE penetration levels ($\sim$17\% solar PV, $\sim$2.3\% battery storage, $\sim$12\% EVs, $\sim$13\% heat pumps). Adoption probabilities are derived from a survey of California residential customers using multilevel regression and poststratification (MRP), with covariates---income, housing type, home size, and climate zone---that map directly to ResStock metadata fields. Buildings are ranked by their predicted adoption probability and assigned technologies from highest to lowest until the target penetration is reached. All solar PV adopters are assigned paired battery storage, reflecting current market trends where approximately 70\% of new California solar installations include batteries \citep{[citation needed]}.

\subsubsection{Load Profile Modification}

For each assigned technology, hourly load profiles are modified as follows:
\begin{itemize}
    \item \textbf{Solar PV}: Each building's PV system is sized to offset 80\% of its annual electricity consumption, subject to a 2--12~kW$_{\text{DC}}$ range reflecting typical residential roof constraints. Under NEM~3.0's asymmetric compensation---where exports earn approximately \$0.05--0.06/kWh versus retail import rates of \$0.45--0.60/kWh---sizing below 100\% offset minimizes low-value exports and optimizes payback periods. Industry analyses find that solar-only systems sized to 60--80\% offset achieve 1--3 year shorter payback periods than 100\%-offset systems under the Net Billing Tariff \citep{AuroraSolar_NEM3_Sizing}, consistent with earlier findings that marginal bill savings per kWh decline as PV-to-load ratios increase under net metering \citep{Darghouth2011}. For systems with paired battery storage (as in our model), the optimal offset is higher because stored energy retains its full retail displacement value; our 80\% target reflects this intermediate case. Hourly generation profiles are simulated using pvlib \citep{pvlib} with PVGIS TMY irradiance data for a single San Diego location (32.9\textdegree N, 117.1\textdegree W). Net consumption is $Q_{i,h}^{\text{net}} = Q_{i,h}^{\text{scaled}} - G_{i,h}^{\text{PV}}$, with export hours compensated at hourly EEC rates under the Net Billing Tariff.
    \item \textbf{Battery storage}: A 13.5~kWh / 5~kW system (Tesla Powerwall equivalent) is dispatched via linear programming to minimize the customer's TOU bill, subject to capacity, power, and round-trip efficiency (90\%) constraints.
    \item \textbf{Electric vehicles}: A daily charging load of 10~kWh ($\sim$30 miles/day) is added during off-peak hours (10~PM--3~AM), reflecting the default TOU-optimized charging schedule.
    \item \textbf{Heat pumps}: Supplementally-assigned buildings have their heating load converted from gas furnace to heat pump characteristics, with COP varying by outdoor temperature.
\end{itemize}

The complete methodology for technology adoption assignment, including the MRP model specification, covariate mapping, and supplemental assignment algorithm, is described in the Supplementary Methods (Section~\ref{sec:tech_adoption}).

\input{technology_adoption_methods}

\subsection{Computational Pipeline}
\label{sec:pipeline}

The analysis is implemented as an end-to-end pipeline that processes all 4,898 SDGE buildings through seven stages: (1)~rate scenario generation, (2)~baseline bill calculation under all scenarios, (3)~technology adoption assignment, (4)~solar profile generation with pvlib, (5)~battery dispatch optimization via linear programming, (6)~post-adoption bill calculation under the Net Billing Tariff, and (7)~distributional analysis and summary statistics.

Bill calculation uses an \textit{anchored billing} approach that ties all designed scenarios to the actual SDGE TOU-DR tariff. First, each building's actual TOU-DR bill $V_i$ is computed using the full tariff calculator, which applies posted TOU rates, subtracts baseline credits ($\sim$\$0.11/kWh on within-baseline consumption), and applies CARE discounts (37\% on volumetric charges). This $V_i$ serves as the building's volumetric component---the bill it would pay under current rates with no fixed charges.

For each designed scenario $j$ with target revenue $R_j^{\text{target}}$ and monthly fixed charges $F_{ij}$, we compute a volumetric scaling factor $\alpha_j$ and anchored bills:
\begin{equation}
\alpha_j = \frac{R_0 \cdot (R_j^{\text{target}} / R^{\text{filed}}) - \sum_i w_i F_{ij}}{R_0}, \qquad
B_{ij} = \alpha_j \cdot V_i + F_{ij}
\end{equation}
where $R_0 = \sum_i w_i V_i$ is the sample's weighted baseline revenue and $R^{\text{filed}}$ is SDGE's filed residential revenue (\$1.562B). This ensures that: (a) the baseline scenario F0\_WF0\_ROE0 ($\alpha = 1$, $F = 0$) produces bills identical to actual TOU-DR; (b) all scenario comparisons preserve the nonlinear tariff structure (baseline credits, CARE discounts); and (c) revenue neutrality holds across scenarios with identical policy parameters.

For post-adoption bills (buildings with solar PV), the anchored approach computes import costs at actual TOU-DR posted rates (including baseline credits and CARE), scales by $\alpha_j$, and subtracts export credits at EEC rates: $B_{ij}^{\text{post}} = \max(\alpha_j \cdot C_{\text{import},i} - C_{\text{export},i},\; 0) + F_{ij}$.

\section{Results}
[Results sections to be filled in]

\subsection{Baseline Rate Comparison}

\subsection{Fixed Charge Impacts}

\subsection{Policy Interventions}

\subsection{Geographic Variation}

\section{Discussion}
\label{sec:discussion}

\subsection{Fixed Charge Regressivity and CPUC Regulatory Context}
\label{sec:regulatory_context}

Our finding that CARE customers experience bill increases under the fixed charge rate structure TOU-DR-F---approximately \$46/year more than under the volumetric TOU-DR rate---while non-CARE customers see only modest increases (\$6/year) is consistent with concerns raised extensively in CPUC Rulemaking R.22-07-005. The mechanism is straightforward: CARE customers have systematically lower consumption than non-CARE customers, so the volumetric rate reduction that accompanies the fixed charge saves them less in absolute terms, while the fixed charge itself (\$6.11/month $= $ \$73/year) represents a larger share of their total bill.

This regressivity persists despite the CARE discount on the fixed charge. The CPUC's adopted structure discounts the fixed charge by approximately 75\% for CARE customers (\$6 vs.\ \$24.15), but this ratio does not match the consumption gap between CARE and non-CARE customers. If CARE customers consume, on average, significantly less than 75\% of non-CARE consumption, the fixed charge discount is insufficient to offset the loss of volumetric savings. Our simulated sample shows a substantial consumption gap between income groups that makes this mechanism operative.

Several parties in R.22-07-005 anticipated this outcome. TURN and NRDC's joint testimony demonstrated that a more steeply graduated fixed charge---\$5/month for CARE, \$42 for middle-income, \$62 for high-income---would produce average CARE savings of \$17--18/month (\$216/year), in stark contrast to the bill increases we observe under the adopted flat structure \citep{NRDC_TURN_01, NRDC_TURN_02}. The Sierra Club proposed an even more progressive design, with zero fixed charges for CARE/FERA customers and \$145/month for upper-income households \citep{SierraClub_R2207005}. Both proposals would have reversed the direction of the distributional impact we document.

The CPUC's E3 rate design tool, which informed Decision D.24-05-028, also showed that bill impacts vary substantially by consumption level and that low-usage customers---who are disproportionately low-income---could experience bill increases under the adopted fixed charge levels \citep{E3_FixedChargeTool}. The CPUC acknowledged this concern but opted for a uniform flat charge (with CARE/FERA discounts) as an interim ``first version'' of the IGFC, citing the need for additional time to develop income verification processes for a more finely graduated structure \citep{CPUC_D2405028}. A working group was established to develop improved income verification for a second-version IGFC with more tiers.

UC Berkeley economist Merideth Fowlie noted that the adopted three-tier structure ``doesn't vary much by income and its promised rate reductions are relatively modest'' \citep{Fowlie2024}. The Renters Alliance highlighted that renting households---which tend to have lower consumption due to smaller dwelling sizes---would be disproportionately harmed, a finding consistent with our results showing that the bill increases from fixed charges are concentrated among low-consumption households regardless of the specific income threshold used for CARE eligibility \citep{RentersAlliance2024}.

Our results thus provide independent, simulation-based confirmation of the distributional concerns raised by consumer advocates in the regulatory proceeding. The quantitative magnitudes we find (\$46/year CARE bill increase vs.\ \$6/year non-CARE increase) are within the range predicted by the various intervenor analyses, lending credibility to both our simulation approach and the advocates' empirical testimony.

\subsection{The Role of the CARE Discount Ratio}
\label{sec:care_discount_ratio}

A striking pattern emerges when comparing the actual TOU-DR-F tariff against our designed fixed charge scenarios (F50, F100). Under TOU-DR-F, CARE customers bear the larger burden: they experience bill increases of \$46/year while non-CARE customers see only \$6/year increases. Under the designed scenarios F50 and F100, this pattern reverses: non-CARE customers face substantially larger bill increases (\$68/year and \$136/year, respectively) while CARE customers see only modest increases (\$4/year and \$8/year).

The explanation lies in the CARE discount ratio $\beta$---the fraction of the non-CARE fixed charge that CARE customers pay. The actual TOU-DR-F tariff sets the CARE fixed charge at \$6.11/month against \$24.60/month for non-CARE, yielding an effective $\beta = 0.248$ (approximately 25\%). Our designed scenarios use $\beta = 0.40$ (40\%), following the CPUC's statutory requirement that the effective CARE discount remain between 30\% and 35\% of non-CARE revenues.\footnote{The CPUC's E3 rate design tool enforces the constraint from Public Utilities Code Section 739.1 that ``the average effective CARE discount shall not be less than 30 percent or more than 35 percent of the revenues that would have been produced for the same billed usage by non-CARE customers.'' Our $\beta = 0.40$ on fixed charges, combined with the 37\% CARE volumetric discount, produces an overall effective discount within this statutory range.}

The divergent outcomes under TOU-DR-F ($\beta = 0.25$) versus the designed scenarios ($\beta = 0.40$) reveal that the CARE discount ratio is a critical---and underappreciated---parameter governing the distributional incidence of fixed charges. At $\beta = 0.25$, the fixed charge discount is insufficient to compensate CARE customers for the volumetric rate reduction they forgo due to their lower consumption. Even though TOU-DR-F's fixed charges are relatively modest (\$6.11/month for CARE), the 75\% discount does not match the consumption gap: CARE customers in our sample consume substantially less than 75\% of non-CARE consumption, so the arithmetic works against them.

At $\beta = 0.40$, the more generous discount shifts the burden toward non-CARE customers, who absorb the full fixed charge while benefiting from volumetric reductions that are proportional to their (higher) consumption. At F50, the non-CARE fixed charge reaches approximately \$37/month, and at F100 approximately \$67/month---far exceeding TOU-DR-F's \$24.60---so the absolute dollar impact on non-CARE customers grows rapidly with the share of T\&D costs recovered through fixed charges.

This finding implies that there exists a break-even discount ratio $\beta^*$ between 0.25 and 0.40 at which CARE and non-CARE customers are affected equally by the introduction of fixed charges. Below $\beta^*$, the fixed charge structure is regressive (CARE customers are disproportionately harmed); above $\beta^*$, the burden shifts to non-CARE customers. The value of $\beta^*$ depends on the consumption ratio between CARE and non-CARE customers and on the level of the fixed charge itself. Our results suggest that the CPUC's adopted discount ratio of approximately 25\% falls below this threshold, consistent with the regressivity concerns raised in R.22-07-005.

\subsection{Extension to Other IOUs}
\label{sec:iou_extension}

Our analysis focuses on SDGE as a case study, with planned extensions to Pacific Gas \& Electric (PG\&E) and Southern California Edison (SCE). The consumption discrepancy documented in Section~\ref{sec:representativeness}---where RASS-calibrated ResStock consumption exceeds utility-reported averages by 46\%---may vary across IOUs depending on the composition of each utility's customer base (e.g., the share of vacant units, partial-year accounts, and multi-family common-area meters). Validating the RASS scaling approach against PG\&E and SCE data will be important for assessing the generalizability of our distributional findings.

The rate design dimensions---fixed charge percentage of T\&D costs, wildfire cost removal, ROE reduction---are generalizable as policy concepts, but the numerical scenarios must be rebuilt for each IOU using their respective filed revenue requirements, TOU rate structures, T\&D cost allocations, rate bases, and customer counts. PG\&E and SCE have different peak/off-peak period definitions and rate ratios, different CARE enrollment rates, and different wildfire cost exposures. We adopt the approach of applying the same policy dimensions with IOU-specific parameters (e.g., ``what does 50\% T\&D fixed charge recovery do for each IOU?''), which is more policy-relevant than imposing identical hypothetical rate structures. The directional findings---that fixed charges are regressive, wildfire cost removal is progressive, and ROE reductions provide broad-based relief---should generalize across utilities, though the magnitudes will differ.

For PV generation profiles, SDGE's compact territory permits a single representative location. PG\&E and SCE span much larger and climatically diverse areas, requiring multiple PV profiles---at minimum one per climate zone---to capture the variation in solar resources from coastal to inland to mountain regions. The pvlib infrastructure and EEC rate data are already prepared for all three IOUs (Section~\ref{sec:eec_data}).

\subsection{Limitations}
\label{sec:limitations}

Several limitations should be noted. First, our analysis uses NREL ResStock simulated building profiles rather than actual utility billing data. While the simulated consumption distributions are calibrated to RASS survey data (Section~\ref{sec:rass}) and reproduce observed patterns across climate zones, housing types, and income groups, they may not perfectly capture the low-usage tail of the CARE customer population. Actual CARE customers may have even lower consumption than our simulated profiles suggest---for example, due to energy-limiting behaviors associated with energy poverty---which would make the regressivity of fixed charges more severe than our estimates indicate.

Second, the consumption gap between our simulated sample mean (5,309~kWh/year) and SDGE's reported average (3,634~kWh/year), discussed in Section~\ref{sec:representativeness}, means that absolute bill levels in our analysis are higher than what the average SDGE customer experiences. However, the distributional analysis depends on the \textit{relative} structure of consumption across customer segments, which is well-preserved in the calibrated sample. The sign and approximate magnitude of bill changes from rate restructuring---the key outputs of our analysis---are robust to level differences in average consumption.

Third, our CARE eligibility assignment is based on ResStock's AMI-based income classifications, which approximate but do not exactly replicate SDGE's CARE enrollment criteria (based on household size and income thresholds, or categorical eligibility through programs like Medi-Cal). The modest discrepancy in CARE rates (25.9\% in our sample vs.\ 28.1\% reported by SDGE) suggests this approximation is reasonable but not exact.

Fourth, our technology adoption assignments use survey-derived probabilities calibrated to aggregate penetration rates. Individual building assignments are probabilistic rather than deterministic, introducing sampling variability in the post-adoption bill distributions. We address this by reporting results at the group level (income terciles, climate zones) where the law of large numbers mitigates individual assignment noise.

Fifth, we use a single solar irradiance profile (San Diego, 32.9\textdegree N) for all buildings in the SDGE territory. This is a reasonable approximation given SDGE's relatively compact service territory spanning three climate zones with similar solar resources. For the planned extension to PG\&E and SCE---whose territories span much larger and climatically diverse areas---multiple PV profiles (at minimum one per climate zone, ideally per-PUMA) will be required to capture geographic variation in solar generation.

\section{Conclusion}
[Conclusion to be filled in]

\section{Acknowledgments}
[To be filled in]

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
